### 2022-07-11 Models.Rmd
summary

---
title: "Summary of models"
output: pdf_document
date: '2022-07-08'
---

Summary of models for seed numbers per plant species

# Richness
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
library(knitr)
library(dplyr)
library(kableExtra)
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") #402 observaciones
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

#Models _ Seed numbers 
# richness
ta_seed<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed<-ta_seed %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed<-filter(ta_seed, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed %>% 
kable(booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed$p.value < 0.05, T, F) )

```

\newpage
visitation rate

```{r, echo=FALSE}
load(".RData")
opts <- options(knitr.kable.NA = "")
knitr::kable(ta_seed2, digits = 4)
```