### 2022-07-14 Models.Rmd
correlation and summary of models

---
title: "Summary of models"
output: pdf_document
date: '2022-07-08'

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

Data base from beefun: plant individuals per sites (13 sites) for five years (2015-2019). Data base has 400 observation, only matching individuals, some with no pollinator frequency (16 observations).
Fruit proportions, seed numbers and seed weight are calculated for each plant individual per plant species in site and year. Richness, visitation rate and frequency of pollinators are also calculated for plant individual per plant species. 
16 plant species but 4 are removed because they are present only one site.

# Correlation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# correlation
a<-Pollinator_Plant %>% select(Plant_gen_sp,fruit_proportion, Seeds,Seed_weight,richness,visitatio_rate,Frequency)
a <- mutate_all(a, ~replace(., is.na(.), 0))

a_cor <- a %>%
  group_by(Plant_gen_sp) %>%
  do(cormat = cor(select(., -matches("Plant_gen_sp"))))

list_1 <- a_cor[[2]]

v <- a_cor[[1]]
names(list_1) <- v

data2 <- do.call(rbind.data.frame, list_1)

corre<-data2 %>%
mutate(newcolum = rownames(data2))%>%
separate(newcolum, into=c('species', 'rows'), extra = "merge", sep = "\\.")%>%
select(species, rows,everything())  %>%
remove_rownames() %>%
  mutate(species = replace(species, duplicated(species), ""))

corre %>% 
   kable(booktabs = TRUE,digits =  4,longtable = T,linesep = c("","","","","","\\addlinespace"),align = "llcccccc") %>%
  kable_styling(font_size = 9,full_width = F, position="left")

```

\newpage

\blandscape

```{r, echo=FALSE, warning=FALSE,message=FALSE, figures-side,fig.height = 13,fig.width=18, out.width="100%",fig.fullwidth=TRUE}

library(tidyverse)
library(broom.mixed)
library(broom)
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)
library(patchwork)
library(ggcorrplot)
library(ggpubr)

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# correlation
a<-Pollinator_Plant %>% select(Plant_gen_sp,fruit_proportion, Seeds,Seed_weight,richness,visitatio_rate,Frequency)
a <- mutate_all(a, ~replace(., is.na(.), 0))

df_with_plots <- a %>%
  group_by(Plant_gen_sp) %>%
  nest() %>%
  mutate(plot = map(data, function(.x) {
    .x  %>%
      cor() %>%
      ggcorrplot::ggcorrplot(show.diag = F, type="lower",lab=TRUE,lab_size=4,tl.cex=12)
  }))

plots1 <- map2(df_with_plots$plot, df_with_plots$Plant_gen_sp, ~(.x + labs(title = .y)))

ggarrange(plots1[[1]], plots1[[2]], plots1[[3]], plots1[[4]],plots1[[5]],plots1[[6]], plots1[[7]], plots1[[8]], plots1[[9]], plots1[[10]],plots1[[11]],plots1[[12]], ncol=4, nrow=3, common.legend = TRUE, legend="right")

```

\elandscape

\clearpage

# Fruit proportion

Summary of models for fruit proportion per plant species. Binomial cbind(fruit_formado, Fruit_No). Site_ID as factor random

## Richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
library(knitr)
library(dplyr)
library(kableExtra)
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

#Models _ fruit proportion 
# richnes
ta_f<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f<-ta_f %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f<-filter(ta_f, !term == "sd__(Intercept)")

ta_f %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# visistation rate
ta_f2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f2<-ta_f2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f2<-filter(ta_f2, !term == "sd__(Intercept)")

ta_f2 %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# frequency
ta_f3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f3<-ta_f3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f3<-filter(ta_f3, !term == "sd__(Intercept)")

ta_f3 %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

# Seed numbers

Summary of models for seed numbers per plant species (LMM). Site_ID as factor random

## Richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") #402 observaciones
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

#Models _ Seed numbers 
# richness
ta_seed<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed<-ta_seed %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed<-filter(ta_seed, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage
## Visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# visitation rate
ta_seed2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed2<-ta_seed2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed2<-filter(ta_seed2, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed2 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")
```

\newpage
## Frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# Frequency
ta_seed3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed3<-ta_seed3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed3<-filter(ta_seed3, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed3 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

# Seed weight

Summary of models for seed weight per plant species (LMM). Site_ID as factor random

## Richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

#Models _ Seed weight 
# richnes
ta_weight<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight<-ta_weight %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight<-filter(ta_weight, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

#Models _ Seed weight 
# visistation rate
ta_weight2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight2<-ta_weight2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight2<-filter(ta_weight2, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight2 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

#Models _ Seed weight 
# frequency
ta_weight3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight3<-ta_weight3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight3<-filter(ta_weight3, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight3 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## join together Lavandula species 
(L.peduncalata (49 obs) and L.stoechas (24 obs))

## Fruit proportion - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richnes
ta_f_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f_la<-ta_f_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f_la<-filter(ta_f_la, !term == "sd__(Intercept)")

ta_f_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_f2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f2_la<-ta_f2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f2_la<-filter(ta_f2_la, !term == "sd__(Intercept)")

ta_f2_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# frequency
ta_f3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f3_la<-ta_f3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f3_la<-filter(ta_f3_la, !term == "sd__(Intercept)")

ta_f3_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# richness
ta_seed_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed_la<-ta_seed_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed_la<-filter(ta_seed_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# visistation rate
ta_seed2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed2_la<-ta_seed2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed2_la<-filter(ta_seed2_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed2_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# Frequency
ta_seed3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed3_la<-ta_seed3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed3_la<-filter(ta_seed3_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed3_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# richnes
ta_weight_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight_la<-ta_weight_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight_la<-filter(ta_weight_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# visistation rate
ta_weight2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight2_la<-ta_weight2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight2_la<-filter(ta_weight2_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight2_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# frequency
ta_weight3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight3_la<-ta_weight3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight3_la<-filter(ta_weight3_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight3_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```