### 2022-07-27 Rmd/Stability.Rmd
stability

---
title: "Stability"
output: pdf_document
date: '2022-07-27'
---

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)
library(dplyr)
library(kableExtra)

Pollinator_Plant<- read.csv("Data/Pollinator_Plant.csv") #400 observaciones
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# fruit proportion stability for each plant species per site
cv_1Fruit<-Pollinator_Plant %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(fruit_proportion),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })

# seed number stability for each plant species per site
cv_1Seed<-Pollinator_Plant %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(Seeds),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })

#visitation rate stability for each plant species per site (all pollinators)
cv_1pollinator<-Pollinator_Plant %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(visitatio_rate),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })

#  pollinator richness for each plant species per site
S<-aggregate(richness ~ Plant_gen_sp+Site_ID, data = Pollinator_Plant,FUN = sum)

# join tables 
full=full_join(cv_1Fruit, cv_1Seed, by = c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., cv_1pollinator, by=c('Plant_gen_sp', 'Site_ID')) %>%
  full_join (., S, by=c('Plant_gen_sp', 'Site_ID'))%>%
  rename(cv_1_fruit=fruit_proportion, cv_1_seed=Seeds, cv_1_visitation=visitatio_rate)

full %>% 
    kable("latex",booktabs = TRUE,longtable = T,linesep = "")
```