### 2022-09-08 Rmd/Stability.Rmd
stability analysis

---
title: "Stability"
output: pdf_document
date: '2022-07-27'

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

For each plant species per site, stability (inverse of the coefficient of variation) of the visitation rate of pollinators (visitation rate calculated like frequency/flower abundance) and also of the fruit proportion and the seed numbers were calculated. Pollinator richness was obtained and furthermore, the log of variance ratio (Lepš et al., 2018) and Loreau & Mazancourt syncrony index were calculated.

Positive values of log var ratio signify synchronization, negative values indicate compensatory dynamics (Lepš et al., 2018).
Loreau & Mazancourt index is standardized between 0 (perfect asynchrony) and 1 (perfect synchrony).

note: NA= one data per year

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)
library(dplyr)
library(kableExtra)
library(reshape2)

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") #400 observaciones
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# fruit proportion stability for each plant species per site
cv_1Fruit<-Pollinator_Plant %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(fruit_proportion),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })%>%
  ungroup()

# seed number stability for each plant species per site
cv_1Seed<-Pollinator_Plant %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(Seeds),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })%>%
  ungroup()

#visitation rate stability for each plant species per site (all pollinators)
cv_1pollinator<-Pollinator_Plant %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(visitatio_rate),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })%>%
  ungroup()

#  pollinator richness for each plant species per site
S<-aggregate(richness ~ Plant_gen_sp+Site_ID, data = Pollinator_Plant,FUN = sum)

## Synchrony ###

focal<- read.csv("../Data/focal.csv") 
focal<-focal[,-1]

focal_0 <- focal[!(focal$Pollinator_gen_sp == "NA NA"),] #eliminar polinizadores NA NA

#Data with species in more than 1 site
focal_0 = focal_0 %>% filter(!Plant_gen_sp %in% spp_out)

focal_0<-subset(focal_0, Plant_gen_sp != 'Ulex australis') #remove ulex
plant_polli<-data.frame(focal_0[,c(1,2,6,29,30)])

df.result <- left_join(Pollinator_Plant, plant_polli, c('Plant_gen_sp', 'Site_ID','Year','plant_id'))
df.result1<-data.frame(df.result[,c(1,2,3,4,25)])
df.result2 <- dcast(df.result, formula = Site_ID + Year +plant_id+Plant_gen_sp ~ Pollinator_gen_sp)

# log variance ratio (Lepš et al., 2018)
#log (Var over the Sum of abundances of all species over time/ Sum of the variance of each species over time)
# Positive values of log var ratio signify synchronization, negative values indicate compensatory dynamics

log_VR <- function(x){
  out <- log((var(rowSums(x))) / (sum(apply(x, MARGIN = 2, FUN = var)))) 
  out                               
}

syn=df.result2%>% 
  group_by(Plant_gen_sp,Site_ID)%>% 
  do(log_VR=log_VR(.[5:129]))

# Loreau & Mazancourt syncrony index
#sum of the tempotral covariances of all species / sum of sqrt of the variances squared
syncLM <- function(x){
  out <- (sum(cov(x))) / (sum(sqrt(diag(cov(x))))^2)
  out
}

syn_2=df.result2%>% 
  group_by(Plant_gen_sp,Site_ID)%>% 
  do(syncLM=syncLM(.[5:129]))

# join tables 
# join tables 
full=full_join(cv_1Fruit, cv_1Seed, by = c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., cv_1pollinator, by=c('Plant_gen_sp', 'Site_ID')) %>%
  full_join (., S, by=c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., syn, by=c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., syn_2, by=c('Plant_gen_sp', 'Site_ID'))%>%
  rename(cv_1_fruit=fruit_proportion, cv_1_seed=Seeds, cv_1_visitation=visitatio_rate)

full$log_VR <- unlist(full$log_VR)
full$syncLM <- unlist(full$syncLM)

full %>% 
    kable("latex",booktabs = TRUE, digits =  4,longtable = T,linesep = "")%>%
  kable_styling(font_size = 8,full_width = F, position="left")
```

\newpage

\blandscape

```{r, echo=FALSE, warning=FALSE,message=FALSE, figures-side,fig.height = 13,fig.width=18, out.width="100%",fig.fullwidth=TRUE}

library(tidyverse)
library(broom.mixed)
library(broom)
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)
library(patchwork)
library(ggcorrplot)
library(ggpubr)

a<-full %>% select(Plant_gen_sp,cv_1_fruit, cv_1_seed,cv_1_visitation,richness,log_VR,syncLM)

## Replace Inf and -Inf with NA
a[is.na(a) | a == "Inf"] <- NA 
a[is.na(a) | a == "-Inf"] <- NA 
a <- mutate_all(a, ~replace(., is.na(.), 0))

df_with_plots <- a %>%
  group_by(Plant_gen_sp) %>%
  nest() %>%
  mutate(plot = map(data, function(.x) {
    .x  %>%
      cor() %>%
      ggcorrplot::ggcorrplot(show.diag = F, type="lower",lab=TRUE,lab_size=4,tl.cex=12)
  }))

plots1 <- map2(df_with_plots$plot, df_with_plots$Plant_gen_sp, ~(.x + labs(title = .y)))

ggarrange(plots1[[1]], plots1[[2]], plots1[[3]], plots1[[4]],plots1[[5]],plots1[[6]], plots1[[7]], plots1[[8]], plots1[[9]], plots1[[10]],plots1[[11]],plots1[[12]], ncol=4, nrow=3, common.legend = TRUE, legend="right")

```

\elandscape

\newpage

Data with species in more than 3 sites. I removed 4 plant species (A.fistulosus, C.libanotis, c.monspeliensis, T.fructicans)

We analysed whether the visitation rate stability is affected by richness and synchrony for each plant species separately.

A) richness + log VR
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
library(tidyr)

sp_out2=full %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id <= 3) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 3 sites
full1 = full %>% filter(!Plant_gen_sp %in% sp_out2)

## Replace Inf and -Inf with NA
full1[is.na(full1) | full1 == "Inf"] <- NA 
full1[is.na(full1) | full1 == "-Inf"] <- NA

sta_polli<-full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

opts <- options(knitr.kable.NA = "")

sta_polli<-sta_polli %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

sta_polli %>% 
    kable("latex",booktabs = TRUE, digits =  4, linesep = c("","","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(sta_polli$p.value < 0.05| is.na(sta_polli$p.value), T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

Goodness of fit measures, p-values for hypothesis tests on residuals, and model convergence information.
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyr)

#goodness of fit measures, p-values for hypothesis tests on residuals, or model convergence information.
sta_polli_glance<-full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table<-sta_polli_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Salvia rosmarinus', 'S.rosmarinus', Plant_gen_sp))

table %>% 
    kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}
library(ggeffects)
library(patchwork)

sta_polli2 <- full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "richness"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)))

p1=sta_polli2$plots[[1]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "C.crispus")+ 
theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_polli2$plots[[2]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_polli2$plots[[3]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=sta_polli2$plots[[4]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_polli2$plots[[5]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_polli2$plots[[6]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_polli2$plots[[7]] + geom_point(data = na.omit( full1 %>% filter(Plant_gen_sp == "Lavandula stoechas")), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
labs(x = "Richness",y="visitation rate stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot (cv_1_visitation~log_VR)
sta_polli2.1 <- full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "log_VR"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)))

p1=sta_polli2.1$plots[[1]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "C.crispus")+
  theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_polli2.1$plots[[2]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_polli2.1$plots[[3]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Cistus salviifolius")), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=sta_polli2.1$plots[[4]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_polli2.1$plots[[5]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Halimium halimifolium")), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_polli2.1$plots[[6]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Lavandula pedunculata")), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_polli2.1$plots[[7]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Lavandula stoechas")), aes(x =  log_VR, y = cv_1_visitation),shape=20)+ 
  labs(x = "log VR",y="visitation rate stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7

```

\newpage

B) richness + Lorea and Mazancourt index
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#model 2: richness + Lorea and Mazancourt index
sta_polli3<-full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+syncLM,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

opts <- options(knitr.kable.NA = "")

sta_polli3<-sta_polli3 %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

sta_polli3 %>% 
    kable("latex",booktabs = TRUE, digits =  4, linesep = c("","","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(sta_polli3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures, p-values for hypothesis tests on residuals, or model convergence information.
sta_polli_glance3<-full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+syncLM,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table2<-sta_polli_glance3 %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Salvia rosmarinus', 'S.rosmarinus', Plant_gen_sp))

table2 %>% 
    kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model2 plot (cv_1_visitation~richness)
sta_polli3.1 <- full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+syncLM,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "richness"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)))

p1=sta_polli3.1$plots[[1]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "C.crispus")+
  theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_polli3.1$plots[[2]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_polli3.1$plots[[3]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=sta_polli3.1$plots[[4]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_polli3.1$plots[[5]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Halimium halimifolium")), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_polli3.1$plots[[6]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_polli3.1$plots[[7]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Lavandula stoechas")), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "Richness",y="visitation rate stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}
# model2 plot (cv_1_visitation~syncLM)
sta_polli3.2 <- full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+syncLM,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "syncLM"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)))

p1=sta_polli3.2$plots[[1]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "C.crispus")+theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_polli3.2$plots[[2]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_polli3.2$plots[[3]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=sta_polli3.2$plots[[4]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_polli3.2$plots[[5]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Halimium halimifolium")), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_polli3.2$plots[[6]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_polli3.2$plots[[7]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Lavandula stoechas")), aes(x =  syncLM, y = cv_1_visitation),shape=20)+ 
  labs(x = "Lorea and Mazancourt index",y="visitation rate stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7
```

\newpage

We analysed the relationship between fruit proportion stability and visitation rate stability
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
# Relationship between fruit proportion stability and visitation rate stability

sta_fruit<-full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_fruit~cv_1_visitation,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

opts <- options(knitr.kable.NA = "")

sta_fruit<-sta_fruit %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

sta_fruit %>% 
    kable("latex",booktabs = TRUE, digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(sta_fruit$p.value < 0.05| is.na(sta_fruit$p.value), T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures, p-values for hypothesis tests on residuals, or model convergence information.
sta_fruit_glance<-full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_fruit~cv_1_visitation,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table3<-sta_fruit_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Salvia rosmarinus', 'S.rosmarinus', Plant_gen_sp))

table3 %>% 
    kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot (fruit proportion stability and visitation rate stability)
sta_fruit.1 <- full1 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_fruit~cv_1_visitation,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "cv_1_visitation"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1)))

p1=sta_fruit.1$plots[[1]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_fruit.1$plots[[2]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Cistus ladanifer")), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_fruit.1$plots[[3]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Cistus salviifolius")), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_fruit.1$plots[[5]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_fruit.1$plots[[6]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Lavandula pedunculata")), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_fruit.1$plots[[7]] + geom_point(data = full1 %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=sta_fruit.1$plots[[8]] + geom_point(data = na.omit(full1 %>% filter(Plant_gen_sp == "Salvia rosmarinus")), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p5+p6+p7+p8
```