### 2022-09-14 Rmd/Stability.Rmd
stability analysis

---
title: "Stability"
output: pdf_document
date: '2022-07-27'

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

For each plant species per site, stability (inverse of the coefficient of variation) of the visitation rate of pollinators (visitation rate calculated like frequency/flower abundance) and also of the fruit proportion and the seed numbers were calculated. Total of pollinator richness (sum) was obtained and furthermore, the log of variance ratio (Lepš et al., 2018) and Loreau & Mazancourt syncrony index were calculated.

Positive values of log var ratio signify synchronization, negative values indicate compensatory dynamics (Lepš et al., 2018).
Loreau & Mazancourt index is standardized between 0 (perfect asynchrony) and 1 (perfect synchrony).

note: NA= one data per year

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)
library(dplyr)
library(kableExtra)
library(reshape2)

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") #400 observaciones
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# datos sin plant_id (mean de las variables porque hay plantas con 2 indv, otras 3, etc)
pol_plant2=Pollinator_Plant %>%
  group_by(Site_ID,Year, Plant_gen_sp) %>%
  summarise_if(is.numeric, mean)

# datos por año y sitio para cada sp de planta (sin plant_id)
cv_1Fruit2<-pol_plant2 %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(fruit_proportion),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })%>%
  ungroup()

# datos por año y sitio para cada sp de planta (sin plant_id)
cv_1Seed2<-pol_plant2 %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(Seeds),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })%>%
  ungroup()

# datos por año y sitio para cada sp de planta (sin plant_id)
cv_1pollinator2<-pol_plant2 %>%
  group_by(Plant_gen_sp,Site_ID) %>%
  summarise_at(vars(visitatio_rate),cv_1 <- function(x, ...) {
    ( mean(x, ...)/sd(x, ...) )
  })%>%
  ungroup()

#suma de la riqueza media por año para cada plant sp y sitio
S2<-aggregate(richness ~ Plant_gen_sp+Site_ID, data = pol_plant2,FUN = sum)

## Synchrony ###

focal<- read.csv("../Data/focal.csv") 
focal<-focal[,-1]

focal_0 <- focal[!(focal$Pollinator_gen_sp == "NA NA"),] #eliminar polinizadores NA NA

#Data with species in more than 1 site
focal_0 = focal_0 %>% filter(!Plant_gen_sp %in% spp_out)

focal_0<-subset(focal_0, Plant_gen_sp != 'Ulex australis') #remove ulex
plant_polli<-data.frame(focal_0[,c(1,2,6,29,30)])

df.result <- left_join(Pollinator_Plant, plant_polli, c('Plant_gen_sp', 'Site_ID','Year','plant_id'))
df.result1<-data.frame(df.result[,c(1,2,3,4,25)])
df.result2 <- dcast(df.result, formula = Site_ID + Year +plant_id+Plant_gen_sp ~ Pollinator_gen_sp)

# sin plant_id
prueba=df.result2 %>%
  group_by(Site_ID,Year, Plant_gen_sp) %>%
  summarise_if(is.numeric, sum)

# log variance ratio (Lepš et al., 2018)
#log (Var over the Sum of abundances of all species over time/ Sum of the variance of each species over time)
# Positive values of log var ratio signify synchronization, negative values indicate compensatory dynamics

log_VR <- function(x){
  out <- log((var(rowSums(x))) / (sum(apply(x, MARGIN = 2, FUN = var)))) 
  out                               
}

# un dato por sp planta, sitio y año (sin plant_id)
syn_2=prueba%>% 
  group_by(Plant_gen_sp,Site_ID)%>% 
  do(log_VR=log_VR(.[4:127]))

# Loreau & Mazancourt syncrony index
#sum of the tempotral covariances of all species / sum of sqrt of the variances squared
syncLM <- function(x){
  out <- (sum(cov(x))) / (sum(sqrt(diag(cov(x))))^2)
  out
}

# un dato por sp planta, sitio y año (sin plant_id)
syn_LM2=prueba%>% 
  group_by(Plant_gen_sp,Site_ID)%>% 
  do(syncLM=syncLM(.[4:127]))

# join tables 
full2=full_join(cv_1Fruit2, cv_1Seed2, by = c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., cv_1pollinator2, by=c('Plant_gen_sp', 'Site_ID')) %>%
  full_join (., S2, by=c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., syn_2, by=c('Plant_gen_sp', 'Site_ID'))%>%
  full_join (., syn_LM2, by=c('Plant_gen_sp', 'Site_ID'))%>%
  rename(cv_1_fruit=fruit_proportion, cv_1_seed=Seeds, cv_1_visitation=visitatio_rate)

full2$log_VR <- unlist(full2$log_VR)
full2$syncLM <- unlist(full2$syncLM)

full2 %>% 
    kable("latex",booktabs = TRUE, digits =  4,longtable = T,linesep = "")%>%
  kable_styling(font_size = 8,full_width = F, position="left")
```

Correlation 
```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

tr2<-full2 %>% select(cv_1_fruit, cv_1_seed,cv_1_visitation,richness)

## Replace Inf and -Inf with NA
tr2[is.na(tr2) | tr2 == "Inf"] <- 0 
tr2[is.na(tr2) | tr2 == "-Inf"] <- 0

tr2 <- mutate_all(tr2, ~replace(., is.na(.), 0))

cr=cor(tr2)

cr %>% 
    kable("latex",booktabs = TRUE, digits =  4)

```

\newpage

\blandscape

```{r, echo=FALSE, warning=FALSE,message=FALSE, figures-side,fig.height = 13,fig.width=18, out.width="100%",fig.fullwidth=TRUE}

library(tidyverse)
library(broom.mixed)
library(broom)
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)
library(patchwork)
library(ggcorrplot)
library(ggpubr)

# correlacion con nuevos calculos (full2 = sin plant_id)
b<-full2 %>% select(Plant_gen_sp,cv_1_fruit, cv_1_seed,cv_1_visitation,richness,log_VR,syncLM)

## Replace Inf and -Inf with NA
b[is.na(b) | b == "Inf"] <- NA 
b[is.na(b) | b == "-Inf"] <- NA

b <- mutate_all(b, ~replace(., is.na(.), 0))

df_with_plotsb <- b %>%
  group_by(Plant_gen_sp) %>%
  nest() %>%
  mutate(plot = map(data, function(.x) {
    .x  %>%
      cor() %>%
      ggcorrplot::ggcorrplot(show.diag = F, type="lower",lab=TRUE)
  }))

plots1 <- map2(df_with_plotsb$plot, df_with_plotsb$Plant_gen_sp, ~(.x + labs(title = .y)))

ggarrange(plots1[[1]], plots1[[2]], plots1[[3]], plots1[[4]],plots1[[5]],plots1[[6]], plots1[[7]], plots1[[8]], plots1[[9]], plots1[[10]],plots1[[11]],plots1[[12]], ncol=4, nrow=3, common.legend = TRUE, legend="right")

```

\elandscape

\newpage

Data with species in more than 3 sites. I removed 4 plant species (A.fistulosus, C.libanotis, c.monspeliensis, T.fructicans)

Salvia rosmarinus = only one year per site (remove)

We analysed the relationship between fruit proportion stability and visitation rate stability

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
# 4 plant species are in 2-3 site (A.fistulosus, C.libanotis, c.monspeliensis, T.fructicans)

# eliminar plant species en new data
sp_out2=full2 %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id <= 3) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 3 sites
full_2 = full2 %>% filter(!Plant_gen_sp %in% sp_out2)

# Salvia rosmarinus solo un dato por año(estabilidad y sincronia =NA). Eliminar

full_2<-subset(full_2, Plant_gen_sp != 'Salvia rosmarinus')

## Replace Inf and -Inf with NA
full_2[is.na(full_2) | full_2 == "Inf"] <- NA 
full_2[is.na(full_2) | full_2 == "-Inf"] <- NA

sta_fruit<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_fruit~cv_1_visitation,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

opts <- options(knitr.kable.NA = "")

sta_fruit<-sta_fruit %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

sta_fruit %>% 
    kable("latex",booktabs = TRUE, digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(sta_fruit$p.value < 0.05| is.na(sta_fruit$p.value), T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")
```

Goodness of fit measures, p-values for hypothesis tests on residuals, and model convergence information.
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures, p-values for hypothesis tests on residuals, or model convergence information.
sta_fruit_glance<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_fruit~cv_1_visitation,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table3<-sta_fruit_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))

table3 %>% 
    kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot (fruit proportion stability and visitation rate stability)
sta_fruit.1 <- full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_fruit~cv_1_visitation,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "cv_1_visitation"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=sta_fruit.1$plots[[1]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_fruit.1$plots[[2]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_fruit.1$plots[[3]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_fruit.1$plots[[5]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_fruit.1$plots[[6]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_fruit.1$plots[[7]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  cv_1_visitation, y = cv_1_fruit),shape=20)+ 
  labs(x = "visitation rate stability",y="fruit proportion stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p5+p6+p7

```

\newpage

We analysed the relationship between seed number stability and visitation rate stability

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
# Relationship between seed number stability and visitation rate stability

sta_seed<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_seed~cv_1_visitation,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

opts <- options(knitr.kable.NA = "")

sta_seed<-sta_seed %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

sta_seed %>% 
kable("latex",booktabs = TRUE, digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(sta_seed$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures

sta_seed_glance<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_seed~cv_1_visitation,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table4<-sta_seed_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Salvia rosmarinus', 'S.rosmarinus', Plant_gen_sp))

table4 %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot (seed number stability and visitation rate stability)
sta_seed.1 <- full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_seed~cv_1_visitation,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "cv_1_visitation"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=sta_seed.1$plots[[1]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "C.crispus")+theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_seed.1$plots[[2]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_seed.1$plots[[3]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=sta_seed.1$plots[[4]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_seed.1$plots[[5]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_seed.1$plots[[6]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=sta_seed.1$plots[[7]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  cv_1_visitation, y = cv_1_seed),shape=20)+ 
  labs(x = "visitation rate stability",y="seed number stability",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7
```

\newpage

Relationship between richness-synchrony with different synchrony indice.

A- Richness ~ log VR
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

rich<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(richness~log_VR,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

rich<-rich %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

rich %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(rich$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures

rich_glance<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(richness~log_VR,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table5<-rich_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))

table5 %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

rich.1 <- full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(richness~log_VR,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "log_VR"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=rich.1$plots[[1]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "C.crispus")+theme_classic()+theme(text = element_text(size=rel(3)))

p2=rich.1$plots[[2]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=rich.1$plots[[3]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=rich.1$plots[[4]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=rich.1$plots[[5]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=rich.1$plots[[6]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=rich.1$plots[[7]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  log_VR, y = richness),shape=20)+ 
  labs(x = "log_VR",y="richness",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7
```

\newpage
A- Richness ~ Loreau & Mazancourt syncrony index
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

rich2<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(richness~syncLM,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

rich2<-rich2 %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

rich2 %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(rich2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures
rich2_glance<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(richness~syncLM,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table6<-rich2_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))

table6 %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

rich.2 <- full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(richness~syncLM,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "syncLM"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=rich.2$plots[[1]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "C.crispus")+theme_classic()+theme(text = element_text(size=rel(3)))

p2=rich.2$plots[[2]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=rich.2$plots[[3]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=rich.2$plots[[4]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=rich.2$plots[[5]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=rich.2$plots[[6]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=rich.2$plots[[7]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  syncLM, y = richness),shape=20)+ 
  labs(x = "syncLM",y="richness",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7
```

\newpage

We analysed whether the visitation rate stability is affected by richness and synchrony for each plant species separately

A) richness + log VR
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

#model 1: richness + log VR
sta_polli<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

sta_polli<-sta_polli %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

sta_polli %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = c("","","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(sta_polli$p.value < 0.05| is.na(sta_polli$p.value), T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

#goodness of fit measures
sta_polli_glance<-full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table7<-sta_polli_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))

table7 %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="left")

```

\newpage
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot (cv_1_visitation~richness)
sta_polli2 <- full_2 %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(cv_1_visitation~richness+log_VR,data))) %>%
  mutate(mod1 = list(ggeffects::ggpredict(mod, terms = "richness"))) %>%
  mutate(plots = list(ggplot(mod1, aes(x, predicted)) + geom_line() +
                        geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=sta_polli2$plots[[1]] + geom_point(data = na.omit(full_2 %>% filter(Plant_gen_sp == "Cistus crispus")), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "richness",y="visitation rate stability",subtitle = "C.crispus")+theme_classic()+theme(text = element_text(size=rel(3)))

p2=sta_polli2$plots[[2]] + geom_point(data = na.omit(full_2 %>% filter(Plant_gen_sp == "Cistus ladanifer")), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "richness",y="visitation rate stability",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=sta_polli2$plots[[3]] + geom_point(data = na.omit (full_2 %>% filter(Plant_gen_sp == "Cistus salviifolius")), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "richness",y="visitation rate stability",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=sta_polli2$plots[[5]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "richness",y="visitation rate stability",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=sta_polli2$plots[[6]] + geom_point(data = full_2 %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = cv_1_visitation),shape=20)+ 
  labs(x = "richness",y="visitation rate stability",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p5+p6
```