### 2022-09-26 Rmd/Models.Rmd
added plots

---
title: "Summary of models"
output: pdf_document
date: '2022-07-08'

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}

---

Data base from beefun: plant individuals per sites (13 sites) for five years (2015-2019). Data base has 400 observation, only matching individuals, some with no pollinator frequency (16 observations).
Fruit proportions, seed numbers and seed weight are calculated for each plant individual per plant species in site and year. Richness, visitation rate and frequency of pollinators are also calculated for plant individual per plant species. 
16 plant species but 4 are removed because they are present only one site.

# Correlation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# correlation
a<-Pollinator_Plant %>% select(Plant_gen_sp,fruit_proportion, Seeds,Seed_weight,richness,visitatio_rate,Frequency)
a <- mutate_all(a, ~replace(., is.na(.), 0))

a_cor <- a %>%
  group_by(Plant_gen_sp) %>%
  do(cormat = cor(select(., -matches("Plant_gen_sp"))))

list_1 <- a_cor[[2]]

v <- a_cor[[1]]
names(list_1) <- v

data2 <- do.call(rbind.data.frame, list_1)

corre<-data2 %>%
mutate(newcolum = rownames(data2))%>%
separate(newcolum, into=c('species', 'rows'), extra = "merge", sep = "\\.")%>%
select(species, rows,everything())  %>%
remove_rownames() %>%
  mutate(species = replace(species, duplicated(species), ""))

corre %>% 
   kable(booktabs = TRUE,digits =  4,longtable = T,linesep = c("","","","","","\\addlinespace"),align = "llcccccc") %>%
  kable_styling(font_size = 9,full_width = F, position="left")

```

\newpage

\blandscape

```{r, echo=FALSE, warning=FALSE,message=FALSE, figures-side,fig.height = 13,fig.width=18, out.width="100%",fig.fullwidth=TRUE}

library(tidyverse)
library(broom.mixed)
library(broom)
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)
library(patchwork)
library(ggcorrplot)
library(ggpubr)

df_with_plots <- a %>%
  group_by(Plant_gen_sp) %>%
  nest() %>%
  mutate(plot = map(data, function(.x) {
    .x  %>%
      cor() %>%
      ggcorrplot::ggcorrplot(show.diag = F, type="lower",lab=TRUE,lab_size=4,tl.cex=12)
  }))

plots1 <- map2(df_with_plots$plot, df_with_plots$Plant_gen_sp, ~(.x + labs(title = .y)))

ggarrange(plots1[[1]], plots1[[2]], plots1[[3]], plots1[[4]],plots1[[5]],plots1[[6]], plots1[[7]], plots1[[8]], plots1[[9]], plots1[[10]],plots1[[11]],plots1[[12]], ncol=4, nrow=3, common.legend = TRUE, legend="right")

```

\elandscape

# Fruit proportion

## Richness

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

# replace NAs by 0
# hay especies de plantas que presentan datos de fruit y seed pero no de polinizador (durante el tiempo observado no se ve ningun polinizador )
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#plot
plots_fruitproportion <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=richness, y=fruit_proportion) + 
   geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_fruitproportion[[1]],plots_fruitproportion[[2]],plots_fruitproportion[[3]],plots_fruitproportion[[4]],plots_fruitproportion[[5]],plots_fruitproportion[[6]],plots_fruitproportion[[7]],plots_fruitproportion[[8]],plots_fruitproportion[[9]],plots_fruitproportion[[10]],plots_fruitproportion[[11]],plots_fruitproportion[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for fruit proportion ~ pollinator richness per plant species. Binomial cbind(fruit_formado, Fruit_No), fruit formada= Fruit yes + fruit preyed + fruit dispersed. Site_ID as factor random

Note: NAs is replaced by zero. The plant species have fruit and seed data but no pollinator visiting were registered. No pollinator visit the plant species during observed time

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ fruit proportion 
# richnes

mod_fruitproportion<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_fruitproportion<-mod_fruitproportion %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_fruitproportion<-filter(mod_fruitproportion, !term == "sd__(Intercept)")

mod_fruitproportion %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(mod_fruitproportion$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_fruitproportion_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_fruitproportion_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot
mod_fruitproportion.1 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "richness"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1= mod_fruitproportion.1$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2= mod_fruitproportion.1$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3= mod_fruitproportion.1$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  richness, y = fruit_proportion),shape=20)+labs(x = "richness",y="fruit proportion",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4= mod_fruitproportion.1$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5= mod_fruitproportion.1$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  richness, y = fruit_proportion),shape=20)+labs(x = "richness",y="fruit proportion",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6= mod_fruitproportion.1$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7= mod_fruitproportion.1$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8= mod_fruitproportion.1$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9= mod_fruitproportion.1$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10= mod_fruitproportion.1$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11= mod_fruitproportion.1$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12= mod_fruitproportion.1$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Visitation rate

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_fruitproportion_2 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=visitatio_rate, y=fruit_proportion) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_fruitproportion_2[[1]],plots_fruitproportion_2[[2]],plots_fruitproportion_2[[3]],plots_fruitproportion_2[[4]],plots_fruitproportion_2[[5]],plots_fruitproportion_2[[6]],plots_fruitproportion_2[[7]],plots_fruitproportion_2[[8]],plots_fruitproportion_2[[9]],plots_fruitproportion_2[[10]],plots_fruitproportion_2[[11]],plots_fruitproportion_2[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for fruit proportion ~ visitation rate per plant species. Binomial cbind(fruit_formado, Fruit_No), fruit formada= Fruit yes + fruit preyed + fruit dispersed. Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

# visistation rate
mod_fruitproportion_2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_fruitproportion_2<-mod_fruitproportion_2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_fruitproportion_2<-filter(mod_fruitproportion_2, !term == "sd__(Intercept)")

mod_fruitproportion_2 %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(mod_fruitproportion_2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_fruitproportion_2_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_fruitproportion_2_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_fruitproportion.2 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "visitatio_rate"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1= mod_fruitproportion.2$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2= mod_fruitproportion.2$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_fruitproportion.2$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4= mod_fruitproportion.2$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5= mod_fruitproportion.2$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6= mod_fruitproportion.2$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7= mod_fruitproportion.2$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8= mod_fruitproportion.2$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9= mod_fruitproportion.2$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10= mod_fruitproportion.2$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_fruitproportion.2$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12= mod_fruitproportion.2$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#Models _ fruit proportion 
# hymenoptera visitation
ta_f4<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Hym_visit + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f4<-ta_f4 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f4<-filter(ta_f4, !term == "sd__(Intercept)")

ta_f4 %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f4$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# frequency
ta_f3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f3<-ta_f3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f3<-filter(ta_f3, !term == "sd__(Intercept)")

ta_f3 %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

# Seed numbers

Summary of models for seed numbers per plant species (LMM). Site_ID as factor random

## Richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") #402 observaciones
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#Models _ Seed numbers 
# richness
ta_seed<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed<-ta_seed %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed<-filter(ta_seed, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage
## Visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# visitation rate
ta_seed2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed2<-ta_seed2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed2<-filter(ta_seed2, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed2 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")
```

\newpage
## Hymenoptera visitation
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# hymenoptera visitation
ta_seed4<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed4<-ta_seed4 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed4<-filter(ta_seed4, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed4 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed4$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage
## Frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

# Frequency
ta_seed3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed3<-ta_seed3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed3<-filter(ta_seed3, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed3 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

# Seed weight

Summary of models for seed weight per plant species (LMM). Site_ID as factor random

## Richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#Models _ Seed weight 
# richnes
ta_weight<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight<-ta_weight %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight<-filter(ta_weight, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#Models _ Seed weight 
# visistation rate
ta_weight2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight2<-ta_weight2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight2<-filter(ta_weight2, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight2 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Hymenoptera visitation
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#Models _ Seed weight 
# Hymenoptera visitation
ta_weight4<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight4<-ta_weight4 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight4<-filter(ta_weight4, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight4 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight4$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#Models _ Seed weight 
# frequency
ta_weight3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight3<-ta_weight3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight3<-filter(ta_weight3, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight3 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## join together Lavandula species 
(L.peduncalata (49 obs) and L.stoechas (24 obs))

## Fruit proportion - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richnes
ta_f_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f_la<-ta_f_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f_la<-filter(ta_f_la, !term == "sd__(Intercept)")

ta_f_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_f2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f2_la<-ta_f2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f2_la<-filter(ta_f2_la, !term == "sd__(Intercept)")

ta_f2_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# hymenoptera visitation
ta_f4_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Hym_visit + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f4_la<-ta_f4_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f4_la<-filter(ta_f4_la, !term == "sd__(Intercept)")

ta_f4_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f4_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# frequency
ta_f3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f3_la<-ta_f3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f3_la<-filter(ta_f3_la, !term == "sd__(Intercept)")

ta_f3_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richness
ta_seed_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed_la<-ta_seed_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed_la<-filter(ta_seed_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_seed2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed2_la<-ta_seed2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed2_la<-filter(ta_seed2_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed2_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# hymenoptera visitation
ta_seed4_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed4_la<-ta_seed4_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed4_la<-filter(ta_seed4_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed4_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed4_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# Frequency
ta_seed3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed3_la<-ta_seed3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed3_la<-filter(ta_seed3_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed3_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richnes
ta_weight_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight_la<-ta_weight_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight_la<-filter(ta_weight_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_weight2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight2_la<-ta_weight2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight2_la<-filter(ta_weight2_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight2_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# hymenoptera visitation
ta_weight4_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight4_la<-ta_weight4_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight4_la<-filter(ta_weight4_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight4_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight4_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# frequency
ta_weight3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight3_la<-ta_weight3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight3_la<-filter(ta_weight3_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight3_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```