### 2022-09-27 Rmd/Models.Rmd
added plots

---
title: "Summary of models"
output: pdf_document
date: '2022-07-08'

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}

---

Data base from beefun: plant individuals per sites (13 sites) for five years (2015-2019). Data base has 400 observation, only matching individuals, some with no pollinator frequency (16 observations).
Fruit proportions, seed numbers and seed weight are calculated for each plant individual per plant species in site and year. Richness, visitation rate and frequency of pollinators are also calculated for plant individual per plant species. 
16 plant species but 4 are removed because they are present only one site.

# Correlation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}
library(tidyverse)
#install.packages("broom.mixed")
library(broom.mixed)
library(broom)
#install.packages("lmerTest")
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# correlation
a<-Pollinator_Plant %>% select(Plant_gen_sp,fruit_proportion, Seeds,Seed_weight,richness,visitatio_rate,Frequency)
a <- mutate_all(a, ~replace(., is.na(.), 0))

a_cor <- a %>%
  group_by(Plant_gen_sp) %>%
  do(cormat = cor(select(., -matches("Plant_gen_sp"))))

list_1 <- a_cor[[2]]

v <- a_cor[[1]]
names(list_1) <- v

data2 <- do.call(rbind.data.frame, list_1)

corre<-data2 %>%
mutate(newcolum = rownames(data2))%>%
separate(newcolum, into=c('species', 'rows'), extra = "merge", sep = "\\.")%>%
select(species, rows,everything())  %>%
remove_rownames() %>%
  mutate(species = replace(species, duplicated(species), ""))

corre %>% 
   kable(booktabs = TRUE,digits =  4,longtable = T,linesep = c("","","","","","\\addlinespace"),align = "llcccccc") %>%
  kable_styling(font_size = 9,full_width = F, position="left")

```

\newpage

\blandscape

```{r, echo=FALSE, warning=FALSE,message=FALSE, figures-side,fig.height = 13,fig.width=18, out.width="100%",fig.fullwidth=TRUE}

library(tidyverse)
library(broom.mixed)
library(broom)
library(lmerTest)
library(dplyr)
library(knitr)
library(kableExtra)
library(patchwork)
library(ggcorrplot)
library(ggpubr)

df_with_plots <- a %>%
  group_by(Plant_gen_sp) %>%
  nest() %>%
  mutate(plot = map(data, function(.x) {
    .x  %>%
      cor() %>%
      ggcorrplot::ggcorrplot(show.diag = F, type="lower",lab=TRUE,lab_size=4,tl.cex=12)
  }))

plots1 <- map2(df_with_plots$plot, df_with_plots$Plant_gen_sp, ~(.x + labs(title = .y)))

ggarrange(plots1[[1]], plots1[[2]], plots1[[3]], plots1[[4]],plots1[[5]],plots1[[6]], plots1[[7]], plots1[[8]], plots1[[9]], plots1[[10]],plots1[[11]],plots1[[12]], ncol=4, nrow=3, common.legend = TRUE, legend="right")

```

\elandscape

## general correlation 
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
# replace NAs by 0
# hay especies de plantas que presentan datos de fruit y seed pero no de polinizador (durante el tiempo observado no se ve ningun polinizador )
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

cor_general<-Pollinator_Plant %>% 
  select(fruit_proportion, Seeds,Seed_weight,richness,visitatio_rate,Frequency)

corr <- round(cor(cor_general), 4)

corr %>% 
    kable("latex",booktabs = TRUE,linesep = "")

```

```{r,,echo=FALSE, warning=FALSE,message=FALSE, out.width="89%"}
ggcorrplot(corr, hc.order = TRUE, type = "lower",
           lab = TRUE)
```

\newpage

# Fruit proportion

## Richness

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

# replace NAs by 0
# hay especies de plantas que presentan datos de fruit y seed pero no de polinizador (durante el tiempo observado no se ve ningun polinizador )
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

#plot
plots_fruitproportion <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=richness, y=fruit_proportion) + 
   geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_fruitproportion[[1]],plots_fruitproportion[[2]],plots_fruitproportion[[3]],plots_fruitproportion[[4]],plots_fruitproportion[[5]],plots_fruitproportion[[6]],plots_fruitproportion[[7]],plots_fruitproportion[[8]],plots_fruitproportion[[9]],plots_fruitproportion[[10]],plots_fruitproportion[[11]],plots_fruitproportion[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for fruit proportion ~ pollinator richness per plant species. Binomial cbind(fruit_formado, Fruit_No), fruit formada= Fruit yes + fruit preyed + fruit dispersed. Site_ID as factor random

Note: NAs is replaced by zero. The plant species have fruit and seed data but no pollinator visiting were registered. No pollinator visit the plant species during observed time

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ fruit proportion 
# richnes

mod_fruitproportion<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_fruitproportion<-mod_fruitproportion %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_fruitproportion<-filter(mod_fruitproportion, !term == "sd__(Intercept)")

mod_fruitproportion %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(mod_fruitproportion$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_fruitproportion_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_fruitproportion_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot
mod_fruitproportion.1 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "richness"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1= mod_fruitproportion.1$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2= mod_fruitproportion.1$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3= mod_fruitproportion.1$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  richness, y = fruit_proportion),shape=20)+labs(x = "richness",y="fruit proportion",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4= mod_fruitproportion.1$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5= mod_fruitproportion.1$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  richness, y = fruit_proportion),shape=20)+labs(x = "richness",y="fruit proportion",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6= mod_fruitproportion.1$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7= mod_fruitproportion.1$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8= mod_fruitproportion.1$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9= mod_fruitproportion.1$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10= mod_fruitproportion.1$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11= mod_fruitproportion.1$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12= mod_fruitproportion.1$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  richness, y = fruit_proportion),shape=20)+ labs(x = "richness",y="fruit proportion",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Visitation rate

visitation rate is calculated like frequency/ flower abundance

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_fruitproportion_2 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=visitatio_rate, y=fruit_proportion) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_fruitproportion_2[[1]],plots_fruitproportion_2[[2]],plots_fruitproportion_2[[3]],plots_fruitproportion_2[[4]],plots_fruitproportion_2[[5]],plots_fruitproportion_2[[6]],plots_fruitproportion_2[[7]],plots_fruitproportion_2[[8]],plots_fruitproportion_2[[9]],plots_fruitproportion_2[[10]],plots_fruitproportion_2[[11]],plots_fruitproportion_2[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for fruit proportion ~ visitation rate per plant species. Binomial cbind(fruit_formado, Fruit_No), fruit formada= Fruit yes + fruit preyed + fruit dispersed. Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

# visistation rate
mod_fruitproportion_2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_fruitproportion_2<-mod_fruitproportion_2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_fruitproportion_2<-filter(mod_fruitproportion_2, !term == "sd__(Intercept)")

mod_fruitproportion_2 %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(mod_fruitproportion_2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_fruitproportion_2_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_fruitproportion_2_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_fruitproportion.2 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "visitatio_rate"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1= mod_fruitproportion.2$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2= mod_fruitproportion.2$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_fruitproportion.2$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4= mod_fruitproportion.2$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5= mod_fruitproportion.2$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6= mod_fruitproportion.2$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7= mod_fruitproportion.2$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8= mod_fruitproportion.2$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9= mod_fruitproportion.2$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10= mod_fruitproportion.2$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_fruitproportion.2$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+ labs(x = "visitation_rate",y="fruit proportion",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12= mod_fruitproportion.2$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  visitatio_rate, y = fruit_proportion),shape=20)+labs(x = "visitation_rate",y="fruit proportion",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Hymenoptera visitation

Hymenoptera visitation is calculated like hymenoptera frequency / Flower abundance
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_fruitproportion_3 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=Hym_visit, y=fruit_proportion) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_fruitproportion_3[[1]],plots_fruitproportion_3[[2]],plots_fruitproportion_3[[3]],plots_fruitproportion_3[[4]],plots_fruitproportion_3[[5]],plots_fruitproportion_3[[6]],plots_fruitproportion_3[[7]],plots_fruitproportion_3[[8]],plots_fruitproportion_3[[9]],plots_fruitproportion_3[[10]],plots_fruitproportion_3[[11]],plots_fruitproportion_3[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for fruit proportion ~ Hymenoptera visitation per plant species. Binomial cbind(fruit_formado, Fruit_No), fruit formada= Fruit yes + fruit preyed + fruit dispersed. Site_ID as factor random
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ fruit proportion 
# hymenoptera visitation
mod_plots_fruitproportion_3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Hym_visit + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_plots_fruitproportion_3<-mod_plots_fruitproportion_3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_plots_fruitproportion_3<-filter(mod_plots_fruitproportion_3, !term == "sd__(Intercept)")

mod_plots_fruitproportion_3 %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(mod_plots_fruitproportion_3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_fruitproportion_3_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Hym_visit + (1|Site_ID),family=binomial,data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_fruitproportion_3_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_fruitproportion.3 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Hym_visit + (1|Site_ID),family=binomial,data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "Hym_visit"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_fruitproportion.3$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_fruitproportion.3$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_fruitproportion.3$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_fruitproportion.3$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_fruitproportion.3$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_fruitproportion.3$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7= mod_fruitproportion.3$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_fruitproportion.3$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_fruitproportion.3$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_fruitproportion.3$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_fruitproportion.3$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_fruitproportion.3$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  Hym_visit, y = fruit_proportion),shape=20)+ labs(x = "Hymenotera_visit",y="fruit proportion",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Pollinator frequency

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_fruitproportion_4 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=Frequency, y=fruit_proportion) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_fruitproportion_4[[1]],plots_fruitproportion_4[[2]],plots_fruitproportion_4[[3]],plots_fruitproportion_4[[4]],plots_fruitproportion_4[[5]],plots_fruitproportion_4[[6]],plots_fruitproportion_4[[7]],plots_fruitproportion_4[[8]],plots_fruitproportion_4[[9]],plots_fruitproportion_4[[10]],plots_fruitproportion_4[[11]],plots_fruitproportion_4[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for fruit proportion ~ pollinator frequency per plant species. Binomial cbind(fruit_formado, Fruit_No), fruit formada= Fruit yes + fruit preyed + fruit dispersed. Site_ID as factor random
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

# frequency
mod_fruitproportion_4<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_fruitproportion_4<-mod_fruitproportion_4 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_fruitproportion_4<-filter(mod_fruitproportion_4, !term == "sd__(Intercept)")

mod_fruitproportion_4 %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(mod_fruitproportion_4$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_fruitproportion_4_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_fruitproportion_4_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_fruitproportion.4 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "Frequency"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
 geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_fruitproportion.4$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_fruitproportion.4$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_fruitproportion.4$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

P4=mod_fruitproportion.4$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_fruitproportion.4$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  Frequency, y = fruit_proportion),shape=20)+labs(x = "Frequency",y="fruit proportion",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_fruitproportion.4$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  Frequency, y = fruit_proportion),shape=20)+labs(x = "Frequency",y="fruit proportion",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_fruitproportion.4$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  Frequency, y = fruit_proportion),shape=20)+labs(x = "Frequency",y="fruit proportion",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8= mod_fruitproportion.4$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  Frequency, y = fruit_proportion),shape=20)+labs(x = "Frequency",y="fruit proportion",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_fruitproportion.4$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_fruitproportion.4$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_fruitproportion.4$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  Frequency, y = fruit_proportion),shape=20)+ labs(x = "Frequency",y="fruit proportion",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_fruitproportion.4$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  Frequency, y = fruit_proportion),shape=20)+  labs(x = "Frequency",y="fruit proportion",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

# Seed number

seed number is calculated like seed number mean per fruit in plant individual per plant species

## Richness

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=richness, y=Seeds) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed[[1]],plots_seed[[2]],plots_seed[[3]],plots_seed[[4]],plots_seed[[5]],plots_seed[[6]],plots_seed[[7]],plots_seed[[8]],plots_seed[[9]],plots_seed[[10]],plots_seed[[11]],plots_seed[[12]], ncol=3, nrow=4)

```

\newpage

Summary of models for seed numbers ~ richness per plant species (LMM). Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ Seed numbers 
# richness
mod_seed<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed<-mod_seed %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed<-filter(mod_seed, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_seed.1 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "richness"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
 geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed.1$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed.1$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed.1$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed.1$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  richness, y = Seeds),shape=20)+labs(x = "richness",y="seed_number",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed.1$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed.1$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  richness, y = Seeds),shape=20)+labs(x = "richness",y="seed_number",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed.1$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  richness, y = Seeds),shape=20)+labs(x = "richness",y="seed_number",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed.1$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  richness, y = Seeds),shape=20)+labs(x = "richness",y="seed_number",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed.1$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed.1$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  richness, y = Seeds),shape=20)+labs(x = "richness",y="seed_number",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed.1$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed.1$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  richness, y = Seeds),shape=20)+ labs(x = "richness",y="seed_number",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Visitation rate

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_2 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=visitatio_rate, y=Seeds) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_2[[1]],plots_seed_2[[2]],plots_seed_2[[3]],plots_seed_2[[4]],plots_seed_2[[5]],plots_seed_2[[6]],plots_seed_2[[7]],plots_seed_2[[8]],plots_seed_2[[9]],plots_seed_2[[10]],plots_seed_2[[11]],plots_seed_2[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed number ~ visitation rate per plant species (LMM). Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

# visitation rate
mod_seed2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed2<-mod_seed2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed2<-filter(mod_seed2, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed2 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")
```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed2_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed2_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_seed.2 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "visitatio_rate")))  %>%  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed.2$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed.2$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed.2$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed.2$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed.2$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed.2$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed.2$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed.2$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  visitatio_rate, y = Seeds),shape=20)+labs(x = "visitation_rate",y="seed_number",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed.2$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  visitatio_rate, y = Seeds),shape=20)+labs(x = "visitation_rate",y="seed_number",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed.2$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11= mod_seed.2$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed.2$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  visitatio_rate, y = Seeds),shape=20)+ labs(x = "visitation_rate",y="seed_number",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Hymenoptera visitation

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_3 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=Hym_visit, y=Seeds) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_3[[1]],plots_seed_3[[2]],plots_seed_3[[3]],plots_seed_3[[4]],plots_seed_3[[5]],plots_seed_3[[6]],plots_seed_3[[7]],plots_seed_3[[8]],plots_seed_3[[9]],plots_seed_3[[10]],plots_seed_3[[11]],plots_seed_3[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed number ~ Hymenoptera visitation per plant species (LMM). Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

# hymenoptera visitation
mod_seed3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed3<-mod_seed3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed3<-filter(mod_seed3, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed3 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed3_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Hym_visit+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed3_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

# model plot 
mod_seed.3 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Hym_visit+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "Hym_visit"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed.3$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  Hym_visit, y = Seeds),shape=20)+  labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed.3$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed.3$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed.3$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed.3$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed.3$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed.3$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed.3$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  Hym_visit, y = Seeds),shape=20)+labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed.3$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed.3$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  Hym_visit, y = Seeds),shape=20)+  labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed.3$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  Hym_visit, y = Seeds),shape=20)+ labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed.3$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  Hym_visit, y = Seeds),shape=20)+  labs(x = "Hymenoptera_visit",y="seed_number",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Pollinator frequency

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_4 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=Frequency, y=Seeds) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_4[[1]],plots_seed_4[[2]],plots_seed_4[[3]],plots_seed_4[[4]],plots_seed_4[[5]],plots_seed_4[[6]],plots_seed_4[[7]],plots_seed_4[[8]],plots_seed_4[[9]],plots_seed_4[[10]],plots_seed_4[[11]],plots_seed_4[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed number ~ pollinator frequency per plant species (LMM). Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

# Frequency
mod_seed4<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed4<-mod_seed4 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed4<-filter(mod_seed4, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed4 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed4$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed4_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed4_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

mod_seed.4 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "Frequency"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed.4$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed.4$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  Frequency, y = Seeds),shape=20)+labs(x = "Frequency",y="seed_number",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed.4$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  Frequency, y = Seeds),shape=20)+labs(x = "Frequency",y="seed_number",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed.4$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed.4$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed.4$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed.4$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed.4$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed.4$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed.4$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed.4$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed.4$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  Frequency, y = Seeds),shape=20)+ labs(x = "Frequency",y="seed_number",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

# Seed weight

seed weight mean per fruit in plant individual per plant species

## Richness

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_weight <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=richness, y=Seed_weight) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_weight[[1]],plots_seed_weight[[2]],plots_seed_weight[[3]],plots_seed_weight[[4]],plots_seed_weight[[5]],plots_seed_weight[[6]],plots_seed_weight[[7]],plots_seed_weight[[8]],plots_seed_weight[[9]],plots_seed_weight[[10]],plots_seed_weight[[11]],plots_seed_weight[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed weight ~ richness per plant species (LMM). Site_ID as factor random
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ Seed weight 
# richnes
mod_seed_weight<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed_weight<-mod_seed_weight %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_weight<-filter(mod_seed_weight, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed_weight %>% 
kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(mod_seed_weight$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed_weight_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed_weight_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

mod_seed_weight.1 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "richness"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed_weight.1$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2= mod_seed_weight.1$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed_weight.1$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed_weight.1$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed_weight.1$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  richness, y = Seed_weight),shape=20)+labs(x = "richness",y="Seed_weight",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed_weight.1$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed_weight.1$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed_weight.1$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed_weight.1$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed_weight.1$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed_weight.1$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  richness, y = Seed_weight),shape=20)+ labs(x = "richness",y="Seed_weight",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed_weight.1$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  richness, y = Seed_weight),shape=20)+  labs(x = "richness",y="Seed_weight",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Visitation rate

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_weight2 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=visitatio_rate, y=Seed_weight) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_weight2[[1]],plots_seed_weight2[[2]],plots_seed_weight2[[3]],plots_seed_weight2[[4]],plots_seed_weight2[[5]],plots_seed_weight2[[6]],plots_seed_weight2[[7]],plots_seed_weight2[[8]],plots_seed_weight2[[9]],plots_seed_weight2[[10]],plots_seed_weight2[[11]],plots_seed_weight2[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed weight ~ visitation rate per plant species (LMM). Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ Seed weight 
# visistation rate
mod_seed_weight2<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed_weight2<-mod_seed_weight2 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_weight2<-filter(mod_seed_weight2, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed_weight2 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed_weight2$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed_weight2_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed_weight2_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

mod_seed_weight.2 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "visitatio_rate"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed_weight.2$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed_weight.2$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed_weight.2$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed_weight.2$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed_weight.2$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed_weight.2$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed_weight.2$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed_weight.2$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+labs(x = "visitation_rate",y="Seed_weight",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed_weight.2$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed_weight.2$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed_weight.2$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed_weight.2$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  visitatio_rate, y = Seed_weight),shape=20)+ labs(x = "visitation_rate",y="Seed_weight",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Hymenoptera visitation

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_weight3 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=Hym_visit, y=Seed_weight) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_weight3[[1]],plots_seed_weight3[[2]],plots_seed_weight3[[3]],plots_seed_weight3[[4]],plots_seed_weight3[[5]],plots_seed_weight3[[6]],plots_seed_weight3[[7]],plots_seed_weight3[[8]],plots_seed_weight3[[9]],plots_seed_weight3[[10]],plots_seed_weight3[[11]],plots_seed_weight3[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed weight ~ hymenoptera visitation per plant species (LMM). Site_ID as factor random
```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ Seed weight 
# Hymenoptera visitation
mod_seed_weight3<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed_weight3<-mod_seed_weight3 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_weight3<-filter(mod_seed_weight3, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed_weight3 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed_weight3$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed_weight3_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Hym_visit+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed_weight3_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

mod_seed_weight.3 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Hym_visit+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "Hym_visit"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed_weight.3$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed_weight.3$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed_weight.3$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed_weight.3$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed_weight.3$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed_weight.3$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed_weight.3$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed_weight.3$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed_weight.3$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed_weight.3$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed_weight.3$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed_weight.3$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  Hym_visit, y = Seed_weight),shape=20)+ labs(x = "Hymenoptera_visit",y="Seed_weight",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## Frequency

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#plot
plots_seed_weight4 <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  group_map(~ ggplot(.) + aes(x=Frequency, y=Seed_weight) + 
              geom_point() + ggtitle(.y[[1]]))

ggarrange(plots_seed_weight4[[1]],plots_seed_weight4[[2]],plots_seed_weight4[[3]],plots_seed_weight4[[4]],plots_seed_weight4[[5]],plots_seed_weight4[[6]],plots_seed_weight4[[7]],plots_seed_weight4[[8]],plots_seed_weight4[[9]],plots_seed_weight4[[10]],plots_seed_weight4[[11]],plots_seed_weight4[[12]], ncol=3,nrow=4)

```

\newpage

Summary of models for seed weight ~ pollinator frequency per plant species (LMM). Site_ID as factor random

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}

#Models _ Seed weight 
# frequency
mod_seed_weight4<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

mod_seed_weight4<-mod_seed_weight4 %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_weight4<-filter(mod_seed_weight4, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_seed_weight4 %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% column_spec(7, bold = ifelse(mod_seed_weight4$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

```{r,echo=FALSE, warning=FALSE,message=FALSE, out.width="80%"}

mod_seed_weight4_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarize(glance(model))%>%
  ungroup()

mod_seed_weight4_glance %>% 
 kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lccccccc")%>%
  kable_styling(font_size = 8,full_width = F, position="center")
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%", fig.align='center'}

mod_seed_weight.4 <- Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  mutate(model1 = list(ggeffects::ggpredict(model, terms = "Frequency"))) %>%
  mutate(plots = list(ggplot(model1, aes(x, predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)))

p1=mod_seed_weight.4$plots[[1]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Asphodelus fistulosus"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "A.fistulosus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p2=mod_seed_weight.4$plots[[2]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus crispus"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "C.crispus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p3=mod_seed_weight.4$plots[[3]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus ladanifer"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "C.ladanifer")+  theme_classic()+theme(text = element_text(size=rel(3)))

p4=mod_seed_weight.4$plots[[4]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus libanotis"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "C.libanotis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p5=mod_seed_weight.4$plots[[5]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus monspeliensis"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "C.monspeliensis")+  theme_classic()+theme(text = element_text(size=rel(3)))

p6=mod_seed_weight.4$plots[[6]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Cistus salviifolius"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "C.salviifolius")+  theme_classic()+theme(text = element_text(size=rel(3)))

p7=mod_seed_weight.4$plots[[7]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium calycinum"), aes(x =  Frequency, y = Seed_weight),shape=20)+ 
  labs(x = "Frequency",y="Seed_weight",subtitle = "H.calycinum")+  theme_classic()+theme(text = element_text(size=rel(3)))

p8=mod_seed_weight.4$plots[[8]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Halimium halimifolium"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "H.halimifolium")+  theme_classic()+theme(text = element_text(size=rel(3)))

p9=mod_seed_weight.4$plots[[9]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula pedunculata"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "L.pedunculata")+  theme_classic()+theme(text = element_text(size=rel(3)))

p10=mod_seed_weight.4$plots[[10]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Lavandula stoechas"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "L.stoechas")+  theme_classic()+theme(text = element_text(size=rel(3)))

p11=mod_seed_weight.4$plots[[11]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Salvia rosmarinus"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "S.rosmarinus")+  theme_classic()+theme(text = element_text(size=rel(3)))

p12=mod_seed_weight.4$plots[[12]] + geom_point(data = Pollinator_Plant %>% filter(Plant_gen_sp == "Teucrium fruticans"), aes(x =  Frequency, y = Seed_weight),shape=20)+ labs(x = "Frequency",y="Seed_weight",subtitle = "T.fruticans")+  theme_classic()+theme(text = element_text(size=rel(3)))

p1+p2+p3+p4+p5+p6+p7+p8+p9+p10+p11+p12

```

\newpage

## join together Lavandula species 
(L.peduncalata (49 obs) and L.stoechas (24 obs))

## Fruit proportion - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richnes
ta_f_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~richness + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f_la<-ta_f_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f_la<-filter(ta_f_la, !term == "sd__(Intercept)")

ta_f_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_f2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~visitatio_rate + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f2_la<-ta_f2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f2_la<-filter(ta_f2_la, !term == "sd__(Intercept)")

ta_f2_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# hymenoptera visitation
ta_f4_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Hym_visit + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f4_la<-ta_f4_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f4_la<-filter(ta_f4_la, !term == "sd__(Intercept)")

ta_f4_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f4_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Fruit proportion - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# frequency
ta_f3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(glmer(cbind(fruit_formado, Fruit_No)~Frequency + (1|Site_ID),family=binomial,data)))%>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_f3_la<-ta_f3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_f3_la<-filter(ta_f3_la, !term == "sd__(Intercept)")

ta_f3_la %>% 
    kable("latex",booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(6, bold = ifelse(ta_f3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richness
ta_seed_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed_la<-ta_seed_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed_la<-filter(ta_seed_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_seed2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed2_la<-ta_seed2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed2_la<-filter(ta_seed2_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed2_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# hymenoptera visitation
ta_seed4_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed4_la<-ta_seed4_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed4_la<-filter(ta_seed4_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed4_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed4_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed number - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# Frequency
ta_seed3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_seed3_la<-ta_seed3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_seed3_la<-filter(ta_seed3_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_seed3_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_seed3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - richness

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# richnes
ta_weight_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~richness+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight_la<-ta_weight_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight_la<-filter(ta_weight_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - visitation rate

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# visistation rate
ta_weight2_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~visitatio_rate+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight2_la<-ta_weight2_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight2_la<-filter(ta_weight2_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight2_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight2_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - hymenoptera visitation

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# hymenoptera visitation
ta_weight4_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Hym_visit+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight4_la<-ta_weight4_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight4_la<-filter(ta_weight4_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight4_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight4_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```

\newpage

## Seed weight - frequency

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="100%"}
Pollinator_Plant<- read.csv("../Data/Pollinator_Plant.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

lavan_tog=Pollinator_Plant %>%
  mutate(Plant_gen_sp = recode(Plant_gen_sp, "Lavandula pedunculata" = "Lavandula","Lavandula stoechas" = "Lavandula"))

# frequency
ta_weight3_la<-lavan_tog %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seed_weight~Frequency+(1|Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

opts <- options(knitr.kable.NA = "")

ta_weight3_la<-ta_weight3_la %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

ta_weight3_la<-filter(ta_weight3_la, !term == "sd__(Intercept)" & !term =="sd__Observation")

ta_weight3_la %>% 
kable("latex", booktabs = TRUE,digits =  4, linesep = c("","\\addlinespace")) %>% 
column_spec(7, bold = ifelse(ta_weight3_la$p.value < 0.05, T, F) )%>%
  kable_styling(position="center", latex_option="hold_position")

```