### 2022-10-06 Rmd/model_with_year.Rmd
models with year

---
title: "Models with year"
output: pdf_document
date: '2022-10-05'
---

# Seed number

**Visitation rate (all pollinators)**

we analyse whether the visitation rate of all pollinators, years and their interaction affect to the seed number for each plant species

**A-** linear model: lm(Seeds~visitatio_rate * Year)

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

library(tidyverse)
library(broom.mixed)
library(broom)
library(lmerTest)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggeffects)
library(patchwork)
library(ggcorrplot)
library(ggpubr)
library(kableExtra)
library(sjPlot)

Pollinator_Plant<- read.csv("../Data/Pollinator_Plant_Apis_separada.csv") 
Pollinator_Plant<-Pollinator_Plant[,-1]

#Species out
spp_out <- Pollinator_Plant %>%
  group_by(Plant_gen_sp) %>%
  summarise(Unique_Id = n_distinct(Site_ID)) %>%
  filter(Unique_Id == 1) %>%
  select(Plant_gen_sp) %>% pull(Plant_gen_sp)

#Data with species in more than 1 site
Pollinator_Plant = Pollinator_Plant %>% filter(!Plant_gen_sp %in% spp_out)

# replace NAs by 0
# hay especies de plantas que presentan datos de fruit y seed pero no de polinizador (durante el tiempo observado no se ve ningun polinizador )
Pollinator_Plant <- mutate_all(Pollinator_Plant, ~replace(., is.na(.), 0))

Pollinator_Plant$Year <- factor(Pollinator_Plant$Year)

#model
mod_seed_year<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds~visitatio_rate * Year,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

mod_seed_year<-mod_seed_year %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_year %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(6, bold = ifelse(mod_seed_year$p.value < 0.05| is.na(mod_seed_year$p.value), T, F) )%>%
  kable_styling(position="left")

```

```{r, echo=FALSE, warning=FALSE,message=FALSE, out.width="80%", fig.align='left'}

mod_seed_year_glance<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds~visitatio_rate * Year,data))) %>%
  summarize(glance(mod))%>%
  ungroup()

table1<-mod_seed_year_glance %>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Asphodelus fistulosus','A.fistulosus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus crispus', 'C.crispus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus ladanifer', 'C.ladanifer', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus libanotis', 'C.libanotis', Plant_gen_sp))%>%
   mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus monspeliensis', 'C.monspeliensis', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Cistus salviifolius', 'C.salviifolius', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium calycinum', 'H. calycinum', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Halimium halimifolium', 'H.halimifolium', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula pedunculata', 'L.pedunculata', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Lavandula stoechas', 'L.stoechas', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Salvia rosmarinus', 'S.rosmarinus', Plant_gen_sp))%>%
  mutate( Plant_gen_sp = if_else(Plant_gen_sp == 'Teucrium fruticans', 'T.fructicans', Plant_gen_sp))

table1 %>% 
    kable("latex",booktabs = TRUE, digits =  3,linesep = "",align = "lcccccccccccc") %>%
  kable_styling(font_size = 7,full_width = F, position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
mod_seed_year.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds~visitatio_rate * Year,data)))

# estimates of model 
ggarrange(plot_model(mod_seed_year.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3), ncol=3,nrow = 4)

```

\newpage
Predicte values of seed number
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
mod_seed_year.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds~visitatio_rate * Year,data)))

ggarrange(plot_model(mod_seed_year.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[5]], type = "int", title = "Cistus monspeliensis",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_seed_year.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE), ncol=3,nrow = 4)

```

*Cistus monspeliensis* - *Halimium halimifolium* - *Teucrium fructicans* 
Warning message:
In predict.lm(model, newdata = fitfram, type = "response", se.fit = se,: prediction from a rank-deficient fit may be misleading

**Importante**: Cistus crispus en 2015 solo tiene dos valores, uno con 0 semillas y visitation rate 0.214; y otro con casi 40 semillas y visitation rate 0

\newpage

Predicted values of Seeds for *Cistus crispus*
```{r,echo=FALSE, warning=FALSE,message=FALSE}
get_model_data(mod_seed_year.1$mod[[2]], type = "int")

```

\newpage

**B-** linear mixed model with site ID like random factor
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

#model
mod_mix_seed_year<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds ~ visitatio_rate * Year + (1 | Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

mod_mix_seed_year<-mod_mix_seed_year %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_mix_seed_year<-filter(mod_mix_seed_year, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_mix_seed_year %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(7, bold = ifelse(mod_mix_seed_year$p.value < 0.05, T, F) )%>%
  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

#model plot
mod_mix_seed_year.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lmer(Seeds ~ visitatio_rate * Year + (1 | Site_ID),data)))

# estimates of model 
ggarrange(plot_model(mod_mix_seed_year.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3), ncol=3,nrow = 4)

```

\newpage
Predicted values of Seeds
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_mix_seed_year.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[4]], type = "int",title = "Cistus libanotis", show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[7]], type = "int",title = " Halimium calycinum", show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_mix_seed_year.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE)
,ncol=3,nrow = 4)

```

*Cistus monspeliensis* - *Halimium halimifolium* - *Teucrium fructicans* 
Could not compute variance-covariance matrix of predictions. No confidence intervals are returned.

**se prueban modelos con year like random slope**

Hay especies que dan problema porque hay menos observaciones que 
random effect. El error dice:  the random-effects parameters and the residual variance (or scale parameter) are probably unidentifiable. Las especies que dan problema y habria que eliminar son: Halimium calycinum, Lavandula pedunculata, Lavandula stoechas, Salvia rosmarinus

\newpage
**Visitation rate of *Apis mellifera* **

**A-** linear model: lm(Seeds ~ Apis_visitatio_rate * Year)
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#model
mod_seed_year_apis<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ Apis_visit_rate * Year,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

mod_seed_year_apis<-mod_seed_year_apis %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_year_apis %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(6, bold = ifelse(mod_seed_year_apis$p.value < 0.05| is.na(mod_seed_year_apis$p.value), T, F) )%>%
  kable_styling(position="left")

```

\newpage

```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
mod_seed_year_apis.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ Apis_visit_rate * Year,data)))

ggarrange(plot_model(mod_seed_year_apis.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_apis.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3), ncol=3,nrow = 4)

```

\newpage

Predicted values of Seeds 
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_seed_year_apis.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_seed_year_apis.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE), ncol=3,nrow = 4)

```

*Asphodelus fistulosus* - *Cistus libanotis*- *Cistus monspeliensis* - *Halimium calycinum* - *Halimium halimifolium* - *Teucrium fructicans*

Warning message: In predict.lm(model, newdata = fitfram, type = "response", se.fit = se,: prediction from a rank-deficient fit may be misleading

\newpage
**B-** linear mixed model with site ID like random factor

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

#model
mod_mix_seed_year_apis<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds ~ Apis_visit_rate * Year + (1 | Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

mod_mix_seed_year_apis<-mod_mix_seed_year_apis %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_mix_seed_year_apis<-filter(mod_mix_seed_year_apis, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_mix_seed_year_apis %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(7, bold = ifelse(mod_mix_seed_year_apis$p.value < 0.05, T, F) )%>%
  kable_styling(position="left")

```

\newpage

```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
#model plot
mod_mix_seed_year_apis.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lmer(Seeds ~ Apis_visit_rate * Year + (1 | Site_ID),data)))

ggarrange(plot_model(mod_mix_seed_year_apis.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_apis.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)
```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_mix_seed_year_apis.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_mix_seed_year_apis.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```

*Asphodelus fistulosus* - *Cistus libanotis*- *Cistus monspeliensis* - *Halimium calycinum* - *Halimium halimifolium* - *Teucrium fructicans*

Could not compute variance-covariance matrix of predictions. No confidence intervals are returned

\newpage
**Visitation rate of pollinator without *Apis mellifera* **

**A-** linear model: lm(Seeds ~ No_Apis_visitatio_rate * Year)
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#model
mod_seed_year_no_apis<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ No_Apis_visit_rate * Year,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

mod_seed_year_no_apis<-mod_seed_year_no_apis %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_year_no_apis %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(6, bold = ifelse(mod_seed_year_no_apis$p.value < 0.05| is.na(mod_seed_year_no_apis$p.value), T, F) )%>%
  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

#model plot
mod_seed_year_no_apis.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ No_Apis_visit_rate * Year,data)))

ggarrange(plot_model(mod_seed_year_no_apis.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_no_apis.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_seed_year_no_apis.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_seed_year_no_apis.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```

*Cistus monspeliensis* - *Halimium halimifolium* - *Teucrium fructicans* 
Warning message:
In predict.lm(model, newdata = fitfram, type = "response", se.fit = se,: prediction from a rank-deficient fit may be misleading

\newpage

**B-** linear mixed model with site ID like random factor

Warning messages:Some predictor variables are on very different scales: consider rescaling. The warning occurred in Salvia rosmarinus

```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

#model
mod_mix_seed_year_no_apis<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds ~ No_Apis_visit_rate * Year + (1 | Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

mod_mix_seed_year_no_apis<-mod_mix_seed_year_no_apis %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_mix_seed_year_no_apis<-filter(mod_mix_seed_year_no_apis, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_mix_seed_year_no_apis %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(7, bold = ifelse(mod_mix_seed_year_no_apis$p.value < 0.05, T, F) )%>%  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

#model plot
mod_mix_seed_year_no_apis.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lmer(Seeds ~ No_Apis_visit_rate * Year + (1 | Site_ID),data)))

ggarrange(plot_model(mod_mix_seed_year_no_apis.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[2]], title = "Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_no_apis.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_mix_seed_year_no_apis.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_mix_seed_year_no_apis.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```

\newpage
**Visitation rate of hymenoptera **

**A-** linear model: lm(Seeds ~ No_Apis_visitatio_rate * Year)
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
#model
mod_seed_year_hym<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ Hym_visit * Year,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

mod_seed_year_hym<-mod_seed_year_hym %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_year_hym %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(6, bold = ifelse(mod_seed_year_hym$p.value < 0.05| is.na(mod_seed_year_hym$p.value), T, F) )%>%
  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

mod_seed_year_hym.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ Hym_visit * Year,data)))

ggarrange(plot_model(mod_seed_year_hym.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_seed_year_hym.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_seed_year_hym.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```

\newpage

**B-** linear mixed model with site ID like random factor
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

#model
mod_mix_seed_year_hym<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds ~ Hym_visit * Year + (1 | Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

mod_mix_seed_year_hym<-mod_mix_seed_year_hym %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_mix_seed_year_hym<-filter(mod_mix_seed_year_hym, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_mix_seed_year_hym %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(7, bold = ifelse(mod_mix_seed_year_hym$p.value < 0.05, T, F) )%>%  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
#model plot
mod_mix_seed_year_hym.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lmer(Seeds ~ Hym_visit * Year + (1 | Site_ID),data)))

ggarrange(plot_model(mod_mix_seed_year_hym.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_mix_seed_year_hym.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_mix_seed_year_hym.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```

\newpage
**Visitation rate of hymenoptera without A.mellifera **

**A-** linear model: lm(Seeds ~ No_Apis_visitatio_rate * Year)
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}
# linear model 
mod_seed_year_hym_no_apis<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ Hym_visit_sin_apis * Year,data))) %>%
  summarize(tidy(mod))%>%
  ungroup()

mod_seed_year_hym_no_apis<-mod_seed_year_hym_no_apis %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_seed_year_hym_no_apis %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(6, bold = ifelse(mod_seed_year_hym_no_apis$p.value < 0.05| is.na(mod_seed_year_hym_no_apis$p.value), T, F) )%>%
  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
#model plot
mod_seed_year_hym_no_apis.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lm(Seeds ~ Hym_visit_sin_apis * Year,data)))

ggarrange(plot_model(mod_seed_year_hym_no_apis.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_seed_year_hym_no_apis.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_seed_year_hym_no_apis.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_seed_year_hym_no_apis.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```

\newpage

**B-** linear mixed model with site ID like random factor
```{r,echo=FALSE, warning=FALSE,message=FALSE,fig.height = 13,fig.width=15, out.width="100%"}

#model
mod_mix_seed_year_hym_no_apis<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(model = list(lmer(Seeds ~ Hym_visit_sin_apis * Year + (1 | Site_ID),data))) %>%
  summarise(broom.mixed::tidy(model)) %>%
  ungroup()

mod_mix_seed_year_hym_no_apis<-mod_mix_seed_year_hym_no_apis %>%
  select(-effect, -group)  %>%
  mutate(Plant_gen_sp = replace(Plant_gen_sp, duplicated(Plant_gen_sp), ""))

mod_mix_seed_year_hym_no_apis<-filter(mod_mix_seed_year_hym_no_apis, !term == "sd__(Intercept)" & !term =="sd__Observation")

mod_mix_seed_year_hym_no_apis %>% 
 kable("latex",booktabs = TRUE, digits =  4,longtable = T, linesep = "") %>% 
column_spec(7, bold = ifelse(mod_mix_seed_year_hym_no_apis$p.value < 0.05, T, F) )%>%  kable_styling(position="left")

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}
#model plot
mod_mix_seed_year_hym_no_apis.1<-Pollinator_Plant %>%
  nest_by(Plant_gen_sp) %>%
  mutate(mod = list(lmer(Seeds ~ Hym_visit_sin_apis * Year + (1 | Site_ID),data)))

ggarrange(plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[1]], title = " Asphodelus fistulosus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[2]], title = "  Cistus crispus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[3]], title = " Cistus ladanifer",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[4]], title = "Cistus libanotis",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[5]], title = "Cistus monspeliensis",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[6]], title = " Cistus salviifolius",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[7]], title = " Halimium calycinum",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[8]], title = " Halimium halimifolium",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[9]], title = " Lavandula pedunculata",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[10]], title = " Lavandula stoechas",show.values = TRUE,value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[11]], title = " Salvia rosmarinus",show.values = TRUE, value.offset = .3),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[12]], title = " Teucrium fruticans",show.values = TRUE, value.offset = .3),ncol=3,nrow = 4)

```

\newpage
```{r,echo=FALSE, warning=FALSE,message=FALSE, fig.height = 20,fig.width=22, out.width="100%"}

ggarrange(plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[1]], type = "int", title = " Asphodelus fistulosus",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[2]], type = "int", title = " Cistus crispus",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[3]], type = "int", title = " Cistus ladanifer",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[4]], type = "int",title = "Cistus libanotis",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[5]], type = "int", title = " Cistus monspeliensis",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[6]], type = "int",title = " Cistus salviifolius",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[7]], type = "int",title = " Halimium calycinum",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[8]], type = "int",title = " Halimium halimifolium",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[9]], type = "int",title = " Lavandula pedunculata",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[10]], type = "int",title = " Lavandula stoechas",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[11]], type = "int",title = " Salvia rosmarinus",show.data = TRUE),
plot_model(mod_mix_seed_year_hym_no_apis.1$mod[[12]], type = "int",title = " Teucrium fruticans",show.data = TRUE),ncol=3,nrow = 4)

```